---
title: "Deseq2"
author: "Valdimar"
date: "2024-03-06"
output: html_document
---
```{r}
library("tximport")
library(DESeq2)
library(TxDb.Mmusculus.UCSC.mm10.ensGene)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(EnsDb.Mmusculus.v79)
library(kableExtra)
library(dplyr)
library(Homo.sapiens)
```


Deseq2 analysis pipeline
```{r}
Deseq2_analysis <- function(metadata) {
  
  # Create an empty list to store results
  result_Deseq2 <- list()
  
  for (i in seq_along(metadata)) {
    
    target_gene <- metadata[[i]]$gene[1]
    clean_gene_name <- sub(" RNA-seq", "", target_gene)
    
    GSE <- metadata[[i]]  # Extract each experiment from the list
    
    # Create the necessary metadata for DESeq2
    samples <- data.frame(sample = GSE$run, condition = GSE$condition, row.names = NULL)
    
    # Create the path to the abundance.h5 file
    path <- file.path("C:", "Users", "valds", "OneDrive - Menntaský", "BS_project", "kallisto_output", GSE$run, "abundance.h5")
    
    #check if files exist
  if (all(file.exists(path))) {
    print("Files exist.")
    if (metadata[[i]][["organism_ch1"]][1]=="Mus musculus"){
      txdb <- TxDb.Mmusculus.UCSC.mm10.ensGene
      k <- keys(txdb, keytype = "TXNAME")
      tx2gene <- AnnotationDbi::select(txdb, k, "GENEID", "TXNAME")
      txi <- tximport(path, type = "kallisto", tx2gene = tx2gene, ignoreTxVersion = TRUE)
    
    ddsTxi <- DESeqDataSetFromTximport(txi,
                                       colData = samples,
                                       design = ~condition)
    
    #recommended step to reduce computational time
    keep <- rowSums(counts(ddsTxi)) >= 10
    dds <- ddsTxi[keep,]
    
    #Sets the reference level to WT
    dds$condition <- relevel(dds$condition, ref = "WT")
    
    #Does the analysis
    dds <- DESeq(dds)
    
    #Saves the results
    res <- results(dds)
    res <- as.data.frame(res)
    
    
    hist_obj <- hist(res$pvalue, breaks = 100, main = "Histogram of p values", xlab = "p value", lwd = 0.3, freq = FALSE, col = "blue")
    
    
    ddscount <- counts(dds, normalized = TRUE)
    colnames(ddscount) <- dds$condition
    
    #Getting gene symbols
    geneIDs1 <- ensembldb::select(EnsDb.Mmusculus.v79, keys = rownames(res), keytype = "GENEID", columns = c("SYMBOL", "GENEID"))
    res$gene <- geneIDs1$SYMBOL[base::match(rownames(res), geneIDs1$GENEID)]
    
    resOrdered <- res[base::order(res$pvalue),]
    resOrdered_with_norm_counts <- merge(as.data.frame(resOrdered), ddscount, by = 'row.names', all = TRUE)
    resOrdered_with_norm_counts <- resOrdered_with_norm_counts[order(resOrdered_with_norm_counts$pvalue),]
    resOrdered_with_norm_counts <- resOrdered_with_norm_counts[resOrdered_with_norm_counts$padj < 0.1,]
    resOrdered_with_norm_counts <- na.omit(resOrdered_with_norm_counts)
    
    all_detected_genes <- na.omit(merge(as.data.frame(resOrdered), ddscount, by = 'row.names', all = TRUE))
    all_detected_genes <- all_detected_genes[order(all_detected_genes$padj),]
    all_detected_genes$Row.names<- all_detected_genes$gene
all_detected_genes$gene <- NULL
names(all_detected_genes)[names(all_detected_genes) == "Row.names"] <- "gene"
all_detected_genes$gene <- toupper(all_detected_genes$gene)
  
    
    #filter out SP1, Cirbp, Rbm3, Rtn3 and Smyd5
    Sp1_Cirbp_Rbm3_Rtn3_Smyd5 <- all_detected_genes[all_detected_genes$gene %in% c("SP1", "CIRBP", "RBM3", "RTN3", "SMYD5", clean_gene_name),]
    


result_Deseq2[[GSE$series_id[1]]] <- list()

# Assign values to the elements of the new list
result_Deseq2[[GSE$series_id[1]]]$MHR <- Sp1_Cirbp_Rbm3_Rtn3_Smyd5
result_Deseq2[[GSE$series_id[1]]]$p_histogram <- hist_obj
result_Deseq2[[GSE$series_id[1]]]$res_summary <- summary(res)
   }
  
   else {TxDb(Homo.sapiens) <- TxDb.Hsapiens.UCSC.hg38.knownGene
        k <- keys(TxDb(Homo.sapiens), keytype = "TXNAME")
        tx2gene <- AnnotationDbi::select(Homo.sapiens, k, "SYMBOL", "TXNAME")
        colnames(tx2gene) <- c("TXNAME", "GENEID")
        tx2gene <- na.omit(tx2gene)
        txi <- tximport(path, type = "kallisto", tx2gene = tx2gene)    
   
        ddsTxi <- DESeqDataSetFromTximport(txi,
                                       colData = samples,
                                       design = ~condition)
      #recommended step to reduce computational time
       keep <- rowSums(counts(ddsTxi)) >= 10
        dds <- ddsTxi[keep,]
    
      #Sets the reference level to WT
        dds$condition <- relevel(dds$condition, ref = "WT")
    
      #Does the analysis
        dds <- DESeq(dds)
    
      #Saves the results
      res <- results(dds)
      res <- as.data.frame(res)
    
      #makes p_value-histogram
      hist_obj <- hist(res$pvalue, breaks = 100, main = "Histogram of p values", xlab = "p value", lwd = 0.3, freq = FALSE, col = "blue")
    
    ddscount <- counts(dds, normalized = TRUE)
    colnames(ddscount) <- dds$condition
    
    #Getting gene symbols
    resOrdered <- res[base::order(res$pvalue),]
    resOrdered_with_norm_counts <- merge(as.data.frame(resOrdered), ddscount, by = 'row.names', all = TRUE)
    resOrdered_with_norm_counts <- resOrdered_with_norm_counts[order(resOrdered_with_norm_counts$pvalue),]
    resOrdered_with_norm_counts <- resOrdered_with_norm_counts[resOrdered_with_norm_counts$padj < 0.1,]
    resOrdered_with_norm_counts <- na.omit(resOrdered_with_norm_counts)
    
    all_detected_genes <- na.omit(merge(as.data.frame(resOrdered), ddscount, by = 'row.names', all = TRUE))
    all_detected_genes <- all_detected_genes[order(all_detected_genes$padj),]
    names(all_detected_genes)[names(all_detected_genes) == "Row.names"] <- "gene"
    all_detected_genes$gene <- toupper(all_detected_genes$gene)
    
    
    #filter out SP1, Cirbp, Rbm3, Rtn3 and Smyd5
    Sp1_Cirbp_Rbm3_Rtn3_Smyd5 <- all_detected_genes[all_detected_genes$gene %in% c("SP1", "CIRBP", "RBM3", "RTN3", "SMYD5", clean_gene_name),]
    
  

result_Deseq2[[GSE$series_id[1]]] <- list()

# Assign values to the elements of the new list
result_Deseq2[[GSE$series_id[1]]]$MHR <- Sp1_Cirbp_Rbm3_Rtn3_Smyd5
result_Deseq2[[GSE$series_id[1]]]$p_histogram <- hist_obj
result_Deseq2[[GSE$series_id[1]]]$res_summary <- summary(res)
     
    }  
    
  }
    else {print("files not found")}
  }

  
  return(result_Deseq2)
}


deseq_result<- Deseq2_analysis(modified_data_list_series_id_fixed)




```





Filter out series whose histogram is obviously not "anti-conservative"

```{r}
# These the following indeces

bad_histogram_indices<- c(4, 10, 13, 18, 20, 21, 22, 31, 33, 36, 38, 40, 54, 57, 58, 59, 63, 66, 67, 85, 94)

#lets keep all "anti-conservative" histograms
deseq_result_good_histo<-deseq_result[-bad_histogram_indices]

```

```{r}
FINAL_RESULT <- list()

for (name in names(deseq_result_good_histo)) {
  FINAL_RESULT[[name]] <- list()  # Initialize the list to store results under the current name
  
  for (i in seq_along(deseq_result[[name]])) {  # Iterate over the indices of the current list
    FINAL_RESULT[[name]]$padj<-deseq_result[[name]]$MHR$padj
    FINAL_RESULT[[name]]$gene <- deseq_result[[name]]$MHR$gene
    FINAL_RESULT[[name]]$log2FoldChange <- deseq_result[[name]]$MHR$log2FoldChange
    target_gene <- modified_data_list[[name]]$gene[1]
    clean_gene_name <- sub(" RNA-seq", "", target_gene)
    #find out how many rows are present
    length <- length(FINAL_RESULT[[name]]$gene)
    # Repeat clean_gene_name 'length' times
    target_gene_vector <- rep(clean_gene_name, length)
    FINAL_RESULT[[name]]$targetgene <- target_gene_vector
  }
}
```

We want to remove every series where the target genes have padj greater than 0.1. This is because we only want instances where there was a significant KD.
```{r}

FINAL_RESULT_KD <- list()

for (i in seq_along(FINAL_RESULT)) {
  indices <- which(FINAL_RESULT[[i]]$targetgene == FINAL_RESULT[[i]]$gene)
  selected_padj <- FINAL_RESULT[[i]]$padj[indices]
  selected_logfold<- FINAL_RESULT[[i]]$log2FoldChange[indices]
  print(selected_padj)
  if (any(selected_padj < 0.1 & selected_logfold < 0 )) {
    FINAL_RESULT_KD[[names(FINAL_RESULT)[i]]] <- FINAL_RESULT[[i]]
  } else {
    print("Not a significant KD/KO")
  }
}

```

Load in Gecko list
```{r}
# Load the readxl package
library(readxl)

# Define the path to your Excel file
excel_file <- "Gecko_list.xlsx"

# Specify the names of the sheets you want to read
sheets_to_read <- c("SP1_repr", "RBM3_repr", "SP1_actv")

# Create an empty list to store the data frames
gecko_list <- list()

# Loop through each sheet and read the data into a data frame
for (sheet_name in sheets_to_read) {
  gecko_list[[sheet_name]] <- read_excel(excel_file, sheet = sheet_name)
}
```


Create the final graphs
```{r}
# Load ggplot2 library
library(ggplot2)

# List of genes of interest
genes_of_interest <- c("RBM3", "SMYD5", "RTN3", "CIRBP", "SP1")


create_gene_plot <- function(exp_name, gene, gene_plots, gecko_list, FINAL_RESULT_KD) {
 
  
  exp_data <- FINAL_RESULT_KD[[exp_name]]

  # Check if the gene is in the experiment data
  if (gene %in% exp_data$gene) {
    index <- which(exp_data$gene == gene)
    log2FoldChange <- exp_data$log2FoldChange[index]
    targetgene <- exp_data$targetgene[index]
    p_value <- exp_data$padj[index]

    # Check if the plot for this gene already exists
    if (gene %in% names(gene_plots)) {
      # Extract the existing plot and add new data to it
      existing_plot <- gene_plots[[gene]]$plot
      new_data <- data.frame(TargetGene = targetgene, Log2FoldChange = log2FoldChange, P_value = p_value)

      # Define the color based on your conditions
      if (gene == "RBM3" && targetgene %in% gecko_list$RBM3_repr$RBM3_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_repr$SP1_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_actv$SP1_actv) {
        label_color <- "blue"
      } else {
        label_color <- "black" # Default color if conditions not met
      }

      
      # Add new data to the plot
      existing_plot <- existing_plot +
        geom_point(data = new_data, aes(x = TargetGene, y = Log2FoldChange), color = label_color, size = ifelse(p_value < 0.1, 3, 2), shape = ifelse(p_value < 0.1, 18, 16)) 

      # Update the plot in the gene_plots list
      gene_plots[[gene]]$plot <- existing_plot
      # Store the exp_data under exp_name
      gene_plots[[gene]]$exp_data[[exp_name]] <- exp_data
    } else {
      # Create a new plot for the gene
      plot_data <- data.frame( TargetGene = targetgene, Log2FoldChange = log2FoldChange, P_value = p_value )

      # Define the color based on your conditions
      if (gene == "RBM3" && targetgene %in% gecko_list$RBM3_repr$RBM3_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_repr$SP1_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_actv$SP1_actv) {
        label_color <- "blue"
      } else {
        label_color <- "black" # Default color if conditions not met
      }

      # Create the plot
      p <- ggplot(plot_data, aes(x = TargetGene, y = Log2FoldChange)) +
        theme_light()+
        geom_point(color = label_color, size = ifelse(p_value < 0.1, 3, 2), shape = ifelse(p_value < 0.1, 18, 16)) +
        geom_hline(yintercept = 0, linetype = "dashed", color = "gray") + 
        geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0, ymax = Inf), fill = "red", alpha = 0.2) +
        geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0), fill = "blue", alpha = 0.2) +
        labs(y = "Log2Fold Change", title = paste("Log2Fold Change for", gene)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank()) + 
        guides(color = FALSE)

      # Store the plot and exp_data under exp_name
gene_plots[[gene]] <- list(plot = p, exp_data = setNames(list(exp_data), exp_name))

    }
  } 

  return(gene_plots) # Ensure to return the updated gene_plots
}

# Create an empty list to store plots for each gene of interest
gene_plots <- list()


# List of genes of interest
genes_of_interest <- c("RBM3", "SMYD5", "RTN3", "CIRBP", "SP1")


# Loop over each experiment in FINAL_RESULT_KD
for (exp_name in names(FINAL_RESULT_KD)) {
  # Add a new target gene to our genes of interest
  new_targetgene <- FINAL_RESULT_KD[[exp_name]]$targetgene[1]
  genes_of_interest <- c(genes_of_interest, new_targetgene)
  
  # Loop over each gene of interest
  for (gene in genes_of_interest) {
    gene_plots <- create_gene_plot(exp_name, gene, gene_plots, gecko_list, FINAL_RESULT_KD)
  }
}


```

Now we want to only keep genes that have at least one significant result (p < 0.1)

```{r}
#first we need to append our first data point into layers in the data frame
gene_plots_tofilter<- gene_plots

for (name in names(gene_plots_tofilter)){
  gene_plots_tofilter[[name]]$plot$layers[[1]]$data <- gene_plots_tofilter[[name]]$plot$data
} 
```


```{r} 
gene_pvalues <- list()

# Get the total number of layers
for (i in seq_along(gene_plots_tofilter)){
  name_geneOFinterest<- names(gene_plots_tofilter[i])

for (j in seq_along(gene_plots_tofilter[[i]]$plot$layers)) { 
  layer <- gene_plots_tofilter[[i]]$plot$layers[[j]]
  gene <- layer$data$TargetGene
  pvalue <- layer$data$P_value
  
  # Append the gene and its corresponding p-value to the list
  gene_pvalues[[name_geneOFinterest]][[j]] <- list(gene = gene, pvalue = pvalue)
}
}
# We do not wish to filter out indexes with NULL values. We would like to keep all indexes to use to filter out data points later.
```

```{r}
gene_pvalues_noNULL<- gene_pvalues
# Filter out NULL values from gene_pvalues and store the result back in gene_pvalues
gene_pvalues_noNULL <- lapply(gene_pvalues_noNULL, function(x) Filter(function(y) !is.null(y$gene) && !is.null(y$pvalue), x))


# Create a new list to store grouped genes and p-values
grouped_gene_pvalues <- list()

for (i in seq_along(gene_pvalues_noNULL)) {
  name_geneOFinterest <- names(gene_pvalues_noNULL)[i]
  
  # Initialize entry for the current gene
  grouped_gene_pvalues[[name_geneOFinterest]] <- list()
  
  # Loop through gene_pvalues_noNULL list
  for (entry in gene_pvalues_noNULL[[i]]) {
    gene <- entry$gene
    pvalue <- entry$pvalue
    
    # Check if gene already exists in grouped_gene_pvalues
    if (gene %in% names(grouped_gene_pvalues[[name_geneOFinterest]])) {
      # Append pvalue to existing gene entry
      grouped_gene_pvalues[[name_geneOFinterest]][[gene]] <- c(grouped_gene_pvalues[[name_geneOFinterest]][[gene]], pvalue)
    } else {
      # Create new entry for gene
      grouped_gene_pvalues[[name_geneOFinterest]][[gene]] <- c(pvalue)
    }
  }
}


# Initialize an empty list to store genes with at least one p-value less than 0.1
filtered_gene_pvalues <- list()

# Loop through each gene in grouped_gene_pvalues
for (gene in names(grouped_gene_pvalues)) {
 
    
    # Initialize an empty list to store p-values less than 0.1 for the current gene
    filtered_gene_pvalues[[gene]] <- list()
    
    # Loop through each sublist of p-values for the current gene
    for (sub_gene in names(grouped_gene_pvalues[[gene]])) {
        any_p_value_below_threshold <- any(grouped_gene_pvalues[[gene]][[sub_gene]] < 0.1)
        
        if (any_p_value_below_threshold) {
            
            filtered_gene_pvalues[[gene]][[sub_gene]] <- grouped_gene_pvalues[[gene]][[sub_gene]]
        } 
    }
}



indices <- list()

# Loop through each column (assuming columns represent genes)
for (name in names(gene_pvalues)) {
  # Initialize an empty vector to store indices for each gene
  indices[[name]]<-c()
  
  # Loop through each gene in the column
  for (i in seq_along(gene_pvalues[[name]])){
    # Check if the gene is present in filtered_gene_pvalues
    if (!is.null(gene_pvalues[[name]][[i]]$gene)){
    if (gene_pvalues[[name]][[i]]$gene %in% names(filtered_gene_pvalues[[name]])) {
      # If yes, store the index
      indices[[name]] <- c(indices[[name]], i)
      }
    }
  }
}


#need to add 2,3,4 since it is important for the graph design
for (name in names(indices)) {
  additional_indices <- c(2, 3, 4)
  indices[[name]] <- union(indices[[name]], additional_indices)
}

```

```{r}
filtered_layers <- list()

for (name in names(indices)) {
  filtered_layers[[name]] <- list()  # Initialize list for each gene
  
  for (i in indices[[name]]) {
    # Check if the index is within the bounds of the layers list
    if (i <= length(gene_plots_tofilter[[name]]$plot$layers)) {
      # Add the layer to the filtered list
      filtered_layers[[name]] <- c(filtered_layers[[name]], gene_plots_tofilter[[name]]$plot$layers[[i]])
    }
  }
}
   

gene_plots_tofilter2<- gene_plots_tofilter



for (name in names(filtered_layers)){
# Filter and assign the layers
gene_plots_tofilter2[[name]]$plot$layers <- filtered_layers[[name]]
}

#we should remove "data" under "plot" because it should be present in layers if its significant, if not then it adds unnecessary x value

GoI<- c("RBM3", "CIRBP", "RTN3", "SMYD5", "SP1")


FINAL_gene_plots<- gene_plots_tofilter2



for (gene_of_interest in GoI) {
  FINAL_gene_plots[[gene_of_interest]]$plot$data <-  FINAL_gene_plots[[gene_of_interest]]$plot$layers[[1]]$data
}
```























NOW do the same but lets not worry about significant knock downs

Create the final graphs

```{r}
# Load ggplot2 library
library(ggplot2)

# List of genes of interest
genes_of_interest <- c("RBM3", "SMYD5", "RTN3", "CIRBP", "SP1")


create_gene_plot <- function(exp_name, gene, gene_plots, gecko_list, FINAL_RESULT) {
 
  
  exp_data <- FINAL_RESULT[[exp_name]]

  # Check if the gene is in the experiment data
  if (gene %in% exp_data$gene) {
    index <- which(exp_data$gene == gene)
    log2FoldChange <- exp_data$log2FoldChange[index]
    targetgene <- exp_data$targetgene[index]
    p_value <- exp_data$padj[index]

    # Check if the plot for this gene already exists
    if (gene %in% names(gene_plots)) {
      # Extract the existing plot and add new data to it
      existing_plot <- gene_plots[[gene]]$plot
      new_data <- data.frame(TargetGene = targetgene, Log2FoldChange = log2FoldChange, P_value = p_value)

      # Define the color based on your conditions
      if (gene == "RBM3" && targetgene %in% gecko_list$RBM3_repr$RBM3_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_repr$SP1_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_actv$SP1_actv) {
        label_color <- "blue"
      } else {
        label_color <- "black" # Default color if conditions not met
      }

      

      
      # Add new data to the plot
      existing_plot <- existing_plot +
        geom_point(data = new_data, aes(x = TargetGene, y = Log2FoldChange), color = label_color, size = ifelse(p_value < 0.1, 3, 2), shape = ifelse(p_value < 0.1, 18, 1)) 

      # Update the plot in the gene_plots list
      gene_plots[[gene]]$plot <- existing_plot
      # Store the exp_data under exp_name
      gene_plots[[gene]]$exp_data[[exp_name]] <- exp_data
    } else {
      # Create a new plot for the gene
      plot_data <- data.frame( TargetGene = targetgene, Log2FoldChange = log2FoldChange, P_value = p_value )

      # Define the color based on your conditions
      if (gene == "RBM3" && targetgene %in% gecko_list$RBM3_repr$RBM3_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_repr$SP1_repr) {
        label_color <- "red"
      } else if (gene == "SP1" && targetgene %in% gecko_list$SP1_actv$SP1_actv) {
        label_color <- "blue"
      } else {
        label_color <- "black" # Default color if conditions not met
      }

      
      # Create the plot
     p <- ggplot(plot_data, aes(x = TargetGene, y = Log2FoldChange))+
        theme_light() +
        geom_point(color = label_color, size = ifelse(p_value < 0.1, 3, 2), shape = ifelse(p_value < 0.1, 18, 1)) +
        geom_hline(yintercept = 0, linetype = "solid", color = "gray") + 
        geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0, ymax = Inf), fill = "red", alpha = 0.2) +
        geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0), fill = "blue", alpha = 0.2) +
        labs(y = "Log2Fold Change", title = paste("Log2Fold Change for", gene)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank()) + 
        guides(color = FALSE)
      

      # Store the plot and exp_data under exp_name
gene_plots[[gene]] <- list(plot = p, exp_data = setNames(list(exp_data), exp_name))

    }
  } 

  return(gene_plots) # Ensure to return the updated gene_plots
}

# Create an empty list to store plots for each gene of interest
gene_plots <- list()


# List of genes of interest
genes_of_interest <- c("RBM3", "SMYD5", "RTN3", "CIRBP", "SP1")


# Loop over each experiment in FINAL_RESULT
for (exp_name in names(FINAL_RESULT)) {
  # Add a new target gene to our genes of interest
  new_targetgene <- FINAL_RESULT[[exp_name]]$targetgene[1]
  genes_of_interest <- c(genes_of_interest, new_targetgene)
  
  # Loop over each gene of interest
  for (gene in genes_of_interest) {
    gene_plots <- create_gene_plot(exp_name, gene, gene_plots, gecko_list, FINAL_RESULT)
  }
}


```

Now we want to only keep genes that have at least one significant result (p < 0.1)

```{r}
#first we need to append our first data point into layers in the data frame
gene_plots_tofilter<- gene_plots

for (name in names(gene_plots_tofilter)){
  gene_plots_tofilter[[name]]$plot$layers[[1]]$data <- gene_plots_tofilter[[name]]$plot$data
} 
```


```{r} 
gene_pvalues <- list()

# Get the total number of layers
for (i in seq_along(gene_plots_tofilter)){
  name_geneOFinterest<- names(gene_plots_tofilter[i])

for (j in seq_along(gene_plots_tofilter[[i]]$plot$layers)) { 
  layer <- gene_plots_tofilter[[i]]$plot$layers[[j]]
  gene <- layer$data$TargetGene
  pvalue <- layer$data$P_value
  
  # Append the gene and its corresponding p-value to the list
  gene_pvalues[[name_geneOFinterest]][[j]] <- list(gene = gene, pvalue = pvalue)
}
}
# We do not wish to filter out indexes with NULL values. We would like to keep all indexes to use to filter out data points later.
```

```{r}
gene_pvalues_noNULL<- gene_pvalues
# Filter out NULL values from gene_pvalues and store the result back in gene_pvalues
gene_pvalues_noNULL <- lapply(gene_pvalues_noNULL, function(x) Filter(function(y) !is.null(y$gene) && !is.null(y$pvalue), x))


# Create a new list to store grouped genes and p-values
grouped_gene_pvalues <- list()

for (i in seq_along(gene_pvalues_noNULL)) {
  name_geneOFinterest <- names(gene_pvalues_noNULL)[i]
  
  # Initialize entry for the current gene
  grouped_gene_pvalues[[name_geneOFinterest]] <- list()
  
  # Loop through gene_pvalues_noNULL list
  for (entry in gene_pvalues_noNULL[[i]]) {
    gene <- entry$gene
    pvalue <- entry$pvalue
    
    # Check if gene already exists in grouped_gene_pvalues
    if (gene %in% names(grouped_gene_pvalues[[name_geneOFinterest]])) {
      # Append pvalue to existing gene entry
      grouped_gene_pvalues[[name_geneOFinterest]][[gene]] <- c(grouped_gene_pvalues[[name_geneOFinterest]][[gene]], pvalue)
    } else {
      # Create new entry for gene
      grouped_gene_pvalues[[name_geneOFinterest]][[gene]] <- c(pvalue)
    }
  }
}


# Initialize an empty list to store genes with at least one p-value less than 0.1
filtered_gene_pvalues <- list()

# Loop through each gene in grouped_gene_pvalues
for (gene in names(grouped_gene_pvalues)) {
 
    
    # Initialize an empty list to store p-values less than 0.1 for the current gene
    filtered_gene_pvalues[[gene]] <- list()
    
    # Loop through each sublist of p-values for the current gene
    for (sub_gene in names(grouped_gene_pvalues[[gene]])) {
        any_p_value_below_threshold <- any(grouped_gene_pvalues[[gene]][[sub_gene]] < 0.1)
        
        if (any_p_value_below_threshold) {
            
            filtered_gene_pvalues[[gene]][[sub_gene]] <- grouped_gene_pvalues[[gene]][[sub_gene]]
        } 
    }
}



indices <- list()

# Loop through each column (assuming columns represent genes)
for (name in names(gene_pvalues)) {
  # Initialize an empty vector to store indices for each gene
  indices[[name]]<-c()
  
  # Loop through each gene in the column
  for (i in seq_along(gene_pvalues[[name]])){
    # Check if the gene is present in filtered_gene_pvalues
    if (!is.null(gene_pvalues[[name]][[i]]$gene)){
    if (gene_pvalues[[name]][[i]]$gene %in% names(filtered_gene_pvalues[[name]])) {
      # If yes, store the index
      indices[[name]] <- c(indices[[name]], i)
      }
    }
  }
}


#need to add 2,3,4 since it is important for the graph design
for (name in names(indices)) {
  additional_indices <- c(2, 3, 4)
  indices[[name]] <- union(indices[[name]], additional_indices)
}

```

```{r}
filtered_layers <- list()

for (name in names(indices)) {
  filtered_layers[[name]] <- list()  # Initialize list for each gene
  
  for (i in indices[[name]]) {
    # Check if the index is within the bounds of the layers list
    if (i <= length(gene_plots_tofilter[[name]]$plot$layers)) {
      # Add the layer to the filtered list
      filtered_layers[[name]] <- c(filtered_layers[[name]], gene_plots_tofilter[[name]]$plot$layers[[i]])
    }
  }
}
   

gene_plots_tofilter2<- gene_plots_tofilter



for (name in names(filtered_layers)){
# Filter and assign the layers
gene_plots_tofilter2[[name]]$plot$layers <- filtered_layers[[name]]
}

#we should remove "data" under "plot" because it should be present in layers if its significant, if not then it adds unnecessary x value

GoI<- c("RBM3", "CIRBP", "RTN3", "SMYD5", "SP1")


FINAL_gene_plots_NOTKD<- gene_plots_tofilter2



for (gene_of_interest in GoI) {
  FINAL_gene_plots_NOTKD[[gene_of_interest]]$plot$data <-  FINAL_gene_plots_NOTKD[[gene_of_interest]]$plot$layers[[1]]$data
}
```

All plots scaled the same
```{r}
FINAL_gene_plots_NOTKD_modified<- FINAL_gene_plots_NOTKD

FINAL_gene_plots_NOTKD_modified[["RBM3"]]$plot<-FINAL_gene_plots_NOTKD_modified[["RBM3"]]$plot+scale_y_continuous(name="Log2Fold breyting", limits=c(-2, 2))

FINAL_gene_plots_NOTKD_modified[["SP1"]]$plot<-FINAL_gene_plots_NOTKD_modified[["SP1"]]$plot+scale_y_continuous(name="Log2Fold breyting", limits=c(-2, 2))

FINAL_gene_plots_NOTKD_modified[["SMYD5"]]$plot<-FINAL_gene_plots_NOTKD_modified[["SMYD5"]]$plot+scale_y_continuous(name="Log2Fold breyting", limits=c(-2, 2))

FINAL_gene_plots_NOTKD_modified[["RTN3"]]$plot<-FINAL_gene_plots_NOTKD_modified[["RTN3"]]$plot+scale_y_continuous(name="Log2Fold breyting", limits=c(-2, 2))

FINAL_gene_plots_NOTKD_modified[["CIRBP"]]$plot<-FINAL_gene_plots_NOTKD_modified[["CIRBP"]]$plot+scale_y_continuous(name="Log2Fold breyting", limits=c(-2, 2))



```



How many series show a DEGs given that there was a significant KO/KD

```{r}

How_many_DEGs_series <- list()

for (i in seq_along(FINAL_RESULT_KD)) {
  if (sum(FINAL_RESULT_KD[[i]]$padj < 0.1) >= 2) {
    series_name <- names(FINAL_RESULT_KD)[i]  # Get the name of the series
    How_many_DEGs_series[[series_name]] <- sum(FINAL_RESULT_KD[[i]]$padj < 0.1)  # Store the count of DEGs
  }
}

```

Let's make a heatmap
```{r}
#Let's start by finding the mean log2fold change for each target gene from each KO/KD of the genes of interest

log2SMYD5 <- list()

for (i in 1:43) {
  log2SMYD5[[i]] <- list(
    targetgene = FINAL_gene_plots_NOTKD$SMYD5$plot$layers[[i]]$data$TargetGene,
    log2foldchange = FINAL_gene_plots_NOTKD$SMYD5$plot$layers[[i]]$data$Log2FoldChange
  )
}



# Initialize an empty list to store log2foldchange values for each target gene
SMYD5log2foldchange_grouped <- list()

# Iterate over each element of the log2RBM3 list
for (i in 1:length(log2SMYD5)) {
  # Extract the target gene name and log2foldchange values
  target_gene <- log2SMYD5[[i]]$targetgene
  log2foldchange <- log2SMYD5[[i]]$log2foldchange
  
  # Check if the target gene already exists in the list
  if (target_gene %in% names(SMYD5log2foldchange_grouped)) {
    # If it exists, append the log2foldchange values to the existing list
    SMYD5log2foldchange_grouped[[target_gene]] <- c(SMYD5log2foldchange_grouped[[target_gene]], log2foldchange)
  } else {
    # If it doesn't exist, create a new entry in the list with the log2foldchange values
    SMYD5log2foldchange_grouped[[target_gene]] <- log2foldchange
  }
}




# Initialize an empty list to store the mean log2foldchange for each target gene
SMYD5mean_log2foldchange <- list()

# Iterate over each target gene in the grouped log2foldchange list
for (gene in names(SMYD5log2foldchange_grouped)) {
  # Calculate the mean log2foldchange for the current target gene
  gene_mean <- mean(SMYD5log2foldchange_grouped[[gene]])
  
  # Store the mean log2foldchange in the result list
  SMYD5mean_log2foldchange[[gene]] <- gene_mean
}

```

```{r}
log2RBM3 <- list()

for (i in 1:42) {
  log2RBM3[[i]] <- list(
    targetgene = FINAL_gene_plots_NOTKD$RBM3$plot$layers[[i]]$data$TargetGene,
    log2foldchange = FINAL_gene_plots_NOTKD$RBM3$plot$layers[[i]]$data$Log2FoldChange
  )
}



# Initialize an empty list to store log2foldchange values for each target gene
RBM3log2foldchange_grouped <- list()

# Iterate over each element of the log2RBM3 list
for (i in 1:length(log2RBM3)) {
  # Extract the target gene name and log2foldchange values
  target_gene <- log2RBM3[[i]]$targetgene
  log2foldchange <- log2RBM3[[i]]$log2foldchange
  
  # Check if the target gene already exists in the list
  if (target_gene %in% names(RBM3log2foldchange_grouped)) {
    # If it exists, append the log2foldchange values to the existing list
    RBM3log2foldchange_grouped[[target_gene]] <- c(RBM3log2foldchange_grouped[[target_gene]], log2foldchange)
  } else {
    # If it doesn't exist, create a new entry in the list with the log2foldchange values
    RBM3log2foldchange_grouped[[target_gene]] <- log2foldchange
  }
}




# Initialize an empty list to store the mean log2foldchange for each target gene
RBM3mean_log2foldchange <- list()

# Iterate over each target gene in the grouped log2foldchange list
for (gene in names(RBM3log2foldchange_grouped)) {
  # Calculate the mean log2foldchange for the current target gene
  gene_mean <- mean(RBM3log2foldchange_grouped[[gene]])
  
  # Store the mean log2foldchange in the result list
  RBM3mean_log2foldchange[[gene]] <- gene_mean
}
```


```{r}
log2RTN3 <- list()

for (i in 1:43) {
  log2RTN3[[i]] <- list(
    targetgene = FINAL_gene_plots_NOTKD$RTN3$plot$layers[[i]]$data$TargetGene,
    log2foldchange = FINAL_gene_plots_NOTKD$RTN3$plot$layers[[i]]$data$Log2FoldChange
  )
}



# Initialize an empty list to store log2foldchange values for each target gene
RTN3log2foldchange_grouped <- list()

# Iterate over each element of the log2RBM3 list
for (i in 1:length(log2RTN3)) {
  # Extract the target gene name and log2foldchange values
  target_gene <- log2RTN3[[i]]$targetgene
  log2foldchange <- log2RTN3[[i]]$log2foldchange
  
  # Check if the target gene already exists in the list
  if (target_gene %in% names(RTN3log2foldchange_grouped)) {
    # If it exists, append the log2foldchange values to the existing list
    RTN3log2foldchange_grouped[[target_gene]] <- c(RTN3log2foldchange_grouped[[target_gene]], log2foldchange)
  } else {
    # If it doesn't exist, create a new entry in the list with the log2foldchange values
    RTN3log2foldchange_grouped[[target_gene]] <- log2foldchange
  }
}




# Initialize an empty list to store the mean log2foldchange for each target gene
RTN3mean_log2foldchange <- list()

# Iterate over each target gene in the grouped log2foldchange list
for (gene in names(RTN3log2foldchange_grouped)) {
  # Calculate the mean log2foldchange for the current target gene
  gene_mean <- mean(RTN3log2foldchange_grouped[[gene]])
  
  # Store the mean log2foldchange in the result list
  RTN3mean_log2foldchange[[gene]] <- gene_mean
}

```

```{r}
log2CIRBP <- list()

for (i in 1:46) {
  log2CIRBP[[i]] <- list(
    targetgene = FINAL_gene_plots_NOTKD$CIRBP$plot$layers[[i]]$data$TargetGene,
    log2foldchange = FINAL_gene_plots_NOTKD$CIRBP$plot$layers[[i]]$data$Log2FoldChange
  )
}



# Initialize an empty list to store log2foldchange values for each target gene
CIRBPlog2foldchange_grouped <- list()

# Iterate over each element of the log2RBM3 list
for (i in 1:length(log2CIRBP)) {
  # Extract the target gene name and log2foldchange values
  target_gene <- log2CIRBP[[i]]$targetgene
  log2foldchange <- log2CIRBP[[i]]$log2foldchange
  
  # Check if the target gene already exists in the list
  if (target_gene %in% names(CIRBPlog2foldchange_grouped)) {
    # If it exists, append the log2foldchange values to the existing list
    CIRBPlog2foldchange_grouped[[target_gene]] <- c(CIRBPlog2foldchange_grouped[[target_gene]], log2foldchange)
  } else {
    # If it doesn't exist, create a new entry in the list with the log2foldchange values
    CIRBPlog2foldchange_grouped[[target_gene]] <- log2foldchange
  }
}




# Initialize an empty list to store the mean log2foldchange for each target gene
CIRBPmean_log2foldchange <- list()

# Iterate over each target gene in the grouped log2foldchange list
for (gene in names(CIRBPlog2foldchange_grouped)) {
  # Calculate the mean log2foldchange for the current target gene
  gene_mean <- mean(CIRBPlog2foldchange_grouped[[gene]])
  
  # Store the mean log2foldchange in the result list
  CIRBPmean_log2foldchange[[gene]] <- gene_mean
}

```


```{r}
log2SP1 <- list()

for (i in 1:31) {
  log2SP1[[i]] <- list(
    targetgene = FINAL_gene_plots_NOTKD$SP1$plot$layers[[i]]$data$TargetGene,
    log2foldchange = FINAL_gene_plots_NOTKD$SP1$plot$layers[[i]]$data$Log2FoldChange
  )
}



# Initialize an empty list to store log2foldchange values for each target gene
SP1log2foldchange_grouped <- list()

# Iterate over each element of the log2RBM3 list
for (i in 1:length(log2SP1)) {
  # Extract the target gene name and log2foldchange values
  target_gene <- log2SP1[[i]]$targetgene
  log2foldchange <- log2SP1[[i]]$log2foldchange
  
  # Check if the target gene already exists in the list
  if (target_gene %in% names(SP1log2foldchange_grouped)) {
    # If it exists, append the log2foldchange values to the existing list
    SP1log2foldchange_grouped[[target_gene]] <- c(SP1log2foldchange_grouped[[target_gene]], log2foldchange)
  } else {
    # If it doesn't exist, create a new entry in the list with the log2foldchange values
    SP1log2foldchange_grouped[[target_gene]] <- log2foldchange
  }
}




# Initialize an empty list to store the mean log2foldchange for each target gene
SP1mean_log2foldchange <- list()

# Iterate over each target gene in the grouped log2foldchange list
for (gene in names(SP1log2foldchange_grouped)) {
  # Calculate the mean log2foldchange for the current target gene
  gene_mean <- mean(SP1log2foldchange_grouped[[gene]])
  
  # Store the mean log2foldchange in the result list
  SP1mean_log2foldchange[[gene]] <- gene_mean
}



```


```{r}
# Define all gene lists
gene_lists <- list(
  RBM3 = RBM3mean_log2foldchange,
  CIRBP = CIRBPmean_log2foldchange,
  SMYD5 = SMYD5mean_log2foldchange,
  SP1 = SP1mean_log2foldchange,
  RTN3 = RTN3mean_log2foldchange
)

#all genes
genenames <- list()
for (i in seq_along(gene_lists)) {
  genenames <- c(genenames, names(gene_lists[[i]]))
}

#only unique names
unique_genenames<-list()
unique_genenames<-unique(genenames)


for (i in unique_genenames){
  for (gene in names(gene_lists)){gene_lists[[gene]]}
  if (i gene_lists)
  }
}


for (gene in names(gene_lists)) {
  if (length(gene_lists[[gene]]) < max_length) {
    gene_lists[[gene]] <- c(gene_lists[[gene]], rep(0, max_length - length(gene_lists[[gene]])))
  }
}

# Create the dataframe
meanlog2foldchange <- data.frame(gene_lists)
rownames(meanlog2foldchange) <- names(RBM3mean_log2foldchange)  # Use the gene names as row names

# Print the dataframe
print(meanlog2foldchange)


```
































Plot the histograms
```{r}


GSE_for_histo <- list()

for (name in names(filtered_gene_pvalues)) {
  for (gene in names(filtered_gene_pvalues[[name]])) {
    for (i in seq_along(gene_plots_tofilter2[[name]]$exp_data)) { 
      if (gene_plots_tofilter2[[name]]$exp_data[[i]]$targetgene[1] == gene) {
        GSE_for_histo[[length(GSE_for_histo) + 1]] <- names(gene_plots_tofilter2[[name]]$exp_data)[i]
      }
    }
  }
}


GSE<-GSE_for_histo[[2]]

# Plot the histogram
plot(deseq_result[[GSE]]$p_histogram , main = paste("Histogram of p-values for", GSE))

```





